---
title: "AMS 317 Linear Regression Project"
author: "Group 3: Hao Yang Lin, Ariadna Sandoya, Arjun Talapatra, Benjamin Novik"
date: "Due Date: November 10, 2024"
output: pdf_document
---
Hypothesis 1: Does **room type** (e.g., private room vs. entire home/apartment) influence the number of **reviews per month**?
Hypothesis 2: Do properties in certain **neighborhoods** receive significantly **more reviews** than others?

```{r}
# Load data and check for missing values
library(readr)
library(dplyr)
library(ggplot2)

data <- read_csv("AB_NYC_2019.csv")

# Summary statistics for variables that are tested for the hypothesis.
summary(data %>% select(room_type, neighbourhood_group, reviews_per_month))
missing_values <- sapply(data, function(x) sum(is.na(x)))
print(missing_values)
```
The resulting output shows us that these following variables have missing values: name, host_name, last_review and reviews_per_month with missing values of 16, 21, 10052 & 10052 respectively. From this, it is shown that the variables, last_review and reviews_per_month are likely correlated. If a listing has no reviews, logically there will not be any reviews for the calculated average. 

```{r}
# Count rows where number_of_reviews is 0 and reviews_per_month is NA
count_na_reviews_per_month <- sum(is.na(data$reviews_per_month) & data$number_of_reviews == 0)

# Count total rows where number_of_reviews is 0
count_zero_reviews <- sum(data$number_of_reviews == 0)

# Check if all rows with number_of_reviews == 0 have reviews_per_month as NA
all_zero_reviews_are_na <- count_na_reviews_per_month == count_zero_reviews
all_zero_reviews_are_na

# Replace NA values in reviews_per_month with 0 where number_of_reviews is 0
data <- data %>%
  mutate(reviews_per_month = ifelse(is.na(reviews_per_month) & number_of_reviews == 0, 0, reviews_per_month))


missing_values <- sapply(data, function(x) sum(is.na(x)))
print(missing_values)
```
As observed after the data cleaning process, the majority of variables do not have any missing values anymore aside from last_review, name and host_name. Because these three variables does not pertain any importance to any interest in our analysis, the values can be left as missing.  

# Data Exploration
```{r}
# Summary statistics for room types
data %>%
  group_by(room_type) %>%
  summarise(
    count = n(),
    avg_price = mean(price, na.rm = TRUE),
    avg_reviews_per_month = mean(reviews_per_month, na.rm = TRUE)
  )

summary_stats <- data %>%
  group_by(room_type) %>%
  summarise(
    mean_reviews = mean(reviews_per_month),
    median_reviews = median(reviews_per_month),
    sd_reviews = sd(reviews_per_month)
  )
summary_stats

# Box plot of Reviews per Month by Room type
ggplot(data, aes(x = room_type, y = reviews_per_month)) +
  geom_boxplot(fill = "lightblue", color = "darkblue") +
  labs(title = "Reviews per Month by Room Type",
       x = "Room Type", y = "Reviews per Month") +
  theme_minimal()

# Distribution of room types
ggplot(data, aes(x = room_type)) +
  geom_bar(fill = "lightgreen") +
  labs(title = "Distribution of Room Types",
       x = "Room Type", y = "Count") +
  theme_minimal()

# Price vs. Reviews per Month
ggplot(data, aes(x = price, y = reviews_per_month)) +
  geom_point(alpha = 0.5, color = "steelblue") +
  labs(title = "Price vs. Reviews per Month",
       x = "Price", y = "Reviews per Month") +
  theme_minimal()

# Distribution of prices
ggplot(data, aes(x = price)) +
  geom_histogram(binwidth = 20, fill = "purple", color = "black") +
  labs(title = "Distribution of Prices",
       x = "Price", y = "Frequency") +
  theme_minimal()
```

# Hypothesis 1: ANOVA Test  
Since our dataset is finitely large, over 40000 values, we can assume normality in our data. Thus, performing anova suffices.

```{r}
# ANOVA test for Reviews per Month by Room Type
anova_reviews <- aov(reviews_per_month ~ room_type, data = data)
summary(anova_reviews)

data <- data %>%
  mutate(log_reviews_per_month = log(reviews_per_month + 0.1))

# Plot the log-transformed data by Room Type
ggplot(data, aes(x = room_type, y = log_reviews_per_month)) +
  geom_boxplot(outlier.colour = "red", outlier.shape = 16, outlier.size = 2) +
  labs(title = "Log-Transformed Reviews per Month by Room Type",
       x = "Room Type", y = "Log of Reviews per Month")

```

# Hypothesis 2: Borough Influence on Reviews per Month
```{r}
# Summary Statistucs of Borough
data %>%
  group_by(neighbourhood_group) %>%
  summarise(
    count = n(),
    avg_reviews_per_month = mean(reviews_per_month, na.rm = TRUE),
    median_reviews_per_month = median(reviews_per_month, na.rm = TRUE)
  )
#  Review per Month by Neighborhood Group
ggplot(data, aes(x = neighbourhood_group, y = reviews_per_month)) +
  geom_boxplot() +
  labs(title = "Reviews per Month by Neighborhood Group", x = "Neighborhood Group", y = "Reviews per Month")

# Kruskal-Wallis test for Reviews per Month by Neighborhood Group
kruskal_test <- kruskal.test(reviews_per_month ~ neighbourhood_group, data = data)
kruskal_test
```